Revisiting the linear theory of hydrodynamic wave modulation

### 2024-06-20 paper/paper.tex
Small additions

\title{Dispersion of short waves riding on longer waves}

@@

TODO extend with background.

@@

Eqs. \ref{eq:dispersion}-\ref{eq:w_L} is the full system of equations that we
exploit in the remainder of this paper.

@@

The L-HS solution requires two approximations in their solution:
First, evaluating the long-wave velocity at $z = 0$
rather than $z = \eta_L$; and second, evaluating the solution to the first order
of the Taylor expansion series.
Both of these approximations cause an underestimate of the short-wave
modulation magnitude (Fig. \ref{fig:wavenumber_modulation}).
The original L-HS solution is a cosine wave with an amplitude equal
to the long-wave steepness.
The other two solutions both introduce an increase in wave modulation
on the crest and a decrease in the trough.
At the crest, the relative modulation predicted by the exact L-HS solution
is 10.7\% larger than that predicted by the 1st order approximate L-HS
solution, while the solution that evaluates the long-wave orbital velocity
at the water surface rather than $z = 0$ is 38.4\% larger than that of the
1st order approximate L-HS solution.

@@

Since wave action is variance divided by the intrinsic frequency $\sigma$,
and variance scales with the square of the amplitude, the modulation of
amplitude is proportional to the square root of the modulation of wave action.

\begin{equation}
\label{eq:wave_amplitude_modulation}
\widetilde{a}(\psi) = a \sqrt{\dfrac{\sigma}{\widetilde{\sigma}} \dfrac{\widetilde{N}}{N}}
\end{equation}
To determine $\widetilde{\sigma}$, we will need to evaluate the modulation of
the gravitational acceleration at the surface of the long wave.

>>>>

%\title{Dispersion of short waves riding on longer waves}
\title{Revisiting the linear theory of hydrodynamic wave modulation}

@@

Working through the theory of \citeA{longuet1960changes,longuet1987propagation}
and the numerical simulation of \citeA{peureux2021unsteady}, I found several
gaps in the current understanding of the hydrodynamic modulation of short waves
based on linear wave theory that I thought were worth revisiting.

@@

Eqs. \ref{eq:dispersion}-\ref{eq:w_L} is the full system of equations that we
exploit in the remainder of this paper.
Note that here we derive the solutions from a different starting set of
equations than that of \citeA{longuet1960changes}.
Rather than starting from the conservation of crests and wave action, as we do
here, they derived their solution from the linear superposition of the velocity
potential of two waves.
It appears the starting from the conservation of crests and wave action allows
for a simpler and more intuitive derivation.

@@

The L-HS solution requires two approximations in their solution:
First, evaluating the long-wave velocity at $z = 0$
rather than $z = \eta_L$; and second, evaluating the solution to the first order
of the Taylor expansion series.
Both of these approximations cause an underestimate of the short-wave
modulation magnitude (Fig. \ref{fig:wavenumber_modulation}).
The original L-HS solution is a cosine wave with an amplitude equal
to the long-wave steepness.
The other two solutions both introduce an increase in wave modulation
on the crest and a decrease in the trough.
At the crest, the relative modulation predicted by the exact L-HS solution
is 10.7\% larger than that predicted by the 1st order approximate L-HS
solution, while the solution that evaluates the long-wave orbital velocity
at the water surface rather than $z = 0$ is 38.4\% larger than that of the
1$^{st}$ order approximate L-HS solution.

@@

Since wave action is variance divided by the intrinsic frequency $\sigma$,
and variance scales with $ga^2$, the modulation of the amplitude is proportional
to the square root of the modulation of wave action.

\begin{equation}
\label{eq:wave_amplitude_modulation}
\widetilde{a}(\psi) = a \sqrt{
  \dfrac{g}{\widetilde{g}}
  \dfrac{\sigma}{\widetilde{\sigma}}
  \dfrac{\widetilde{N}}{N}}
\end{equation}
To determine $\widetilde{\sigma}$, we will need to evaluate the modulation of
the gravitational acceleration at the surface of the long wave.
